A Motion Blur Resilient Fiducial Marker For Quadcopter Imaging

### 2014-06-03 WACV/paper.tex
Enlarged overall_flow.jpg, compressed tables

\usepackage{wacv}
\usepackage{times}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[font=small]{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{color}
\usepackage{url}
\usepackage{comment}
\usepackage{multirow}
\usepackage{tabularx}
\newcolumntype{Y}{>{\centering\arraybackslash}X}
% Include other packages here, before hyperref.

@@

\noindent\textbf{Contribution}~~To address these problems, we propose a
fiducial that is designed to be resistant to motion blur. Our design is based on
concentric circles as shown in Figure \ref{fig:teaser}-(c,d). The design is based
on the observation that motion blur from a quadcopter tends to be linear in
nature. As such, when our code is blurred, there is no blur in the direction
perpendicular to the direction of motion.  This allows the signature of the
fiducial to remain intact in any direction. In addition, by using concentric
rings, we can treat the presence or absence of a ring as a bit, allowing us to
assign a code to the marker. Our experiments show that this design can
significantly outperform existing codes in the presence of motion blur. As far
as we are aware, this is the first work to propose a blur resistant marker.

@@

\begin{figure}
\includegraphics[width=\linewidth]{teaser.pdf}
\caption{This figure shows the experimental setup and comparison of
output of ALVAR~\cite{alvar} (for detecting ARTag) versus our blur resistant
fiducial in a normal and blurred video frame. The scene contains two ARTags and our code.
(a) shows that the ARTag can be detected when the image has no blur.
(b) shows that ARTags are no longer detectable due to the blur.
(c-d) show the same images in (a) and (b) where our code is being detected.}
\label{fig:teaser}
\end{figure}

@@

\noindent{\textbf{Fiducials}}~Figure~\ref{fig:previous_work} shows several
examples of existing fiducial markers.  Many designs use a two dimensional
barcode inside a rectangular grid. One example of such a fiducial is from the
ARToolkit~\cite{ARToolkit02}, a well known toolkit used in many augmented
reality (AR) applications. Kato and Billinghurst~\cite{kato-artoolkit}
demonstrated the use of ARToolkit to find the pose in video based AR
conferencing system.

Fiala~\cite{Fiala05} proposed a fiducial termed, ARTag, which is a
bi-tonal system consisting of a square border and an interior 6$\times$6 grid
of black or white cells. The improvement in ARTag compared to ARToolkit was
detection of corners instead of detection of lines to find possible pattern.
This proved to be more efficient than \cite{ARToolkit02} in terms of
recognition rate as well as the number of different patterns which can be
created.   The reliance on both line and corner detection hampers recognition
under motion blur.

@@

\noindent{\textbf{Tracking}}~~Fiducial detection between successive video
frames can be considered as a tracking problem where the tracked object is
fiducial itself.  There is a very large body of research dedicated to tracking and interested
readers are referred to~\cite{Yilmaz:2006} for a good survey.

Most tracking methods (e.g.,~\cite{Ross:2008,Wu:2009,Perez02,Mei:2009}) assume
the image sequence to be blur free. In reality, however, the presence of motion blur in 
a video sequence is often unavoidable. To this end, Wu et al.~\cite{Wu:2011}
proposed the Blur-driven Tracker (BLUT) framework for tracking motion-blurred targets. BLUT
is based on the observation that although motion blurs degrade the visual
features of the target, they at the same time, provide useful cues about the
movements to help tracking.

The BLUT framework successfully tracks blurred target when there is uniform motion
and the position of tracked object does not change drastically in successive
frames. But as we will demonstrate in Section 4, the erratic motion from quadcopters as
well as the problem of dropped video frames is beyond the current ability of such trackers.

@@

Before we detail the design of our fiducial, we examine the performance of two
popular fiducials under motion blur.  Specifically we examine ARTags
~\cite{Fiala05} and PiTags~\cite{Pitag13} given their difference in geometry
design and the availability of API to develop applications to recognize the tags.
RUNE-tag~\cite{runetag11} currently does not provide access to the an
implementation to generate or recognize the tag.  The circular data
matrix~\cite{NaimarkF02} is available as a commercial product, however it
requires proprietary hardware to recognize the fiducial.

@@

\begin{figure}
\centering
  \includegraphics[width=.18\linewidth]{newconcentric_00.pdf}
  \includegraphics[width=.18\linewidth]{newconcentric_01.pdf}
  \includegraphics[width=.18\linewidth]{newconcentric_10.pdf}
  \includegraphics[width=.18\linewidth]{newconcentric_11.pdf}
  \caption{Two bit binary coded fiducials (from left to right: binary code 00,
  binary code 01, binary code 10, binary code 11.)}
  \label{fig:fiducials}
\end{figure}

Depending on which ring is present or absent, the resulting binary code will
change. The number of different patterns depends on the number of bits in the
binary code. For example, if the binary code has three bits, there will be a
maximum of three rings between marker-rings and we end up with eight
different patterns.  Figure \ref{fig:fiducials} shows a two bit binary coded
fiducials.

@@

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{blur_direction.pdf}
\caption{Figure showing how change in blur direction changes the location of
unblurred linear pattern.}
\label{fig:blur_direction}
\end{figure}

@@

\begin{figure}[ht!]
\includegraphics[width=\linewidth]{overall_flow.pdf}
\caption{An overview of our algorithm to detect and classify fiducial.
The four step process includes (Step 1) Gabor filtering,
(Step 2) component clustering, (Step 3) PCA to determine the dominant direction,
and (Step 4) K-NN classification using training-data captured for each pattern.}
\label{fig:overall_flow}
\end{figure}

@@

After finding the intensity profile, we examine it by
projecting the pixel intensities onto the PCA direction and record the
number and width of transitions.  If the number of transitions or the
width of the transitions is not consistent with what is allowable by
our code, we reject the clustered region as a potential fiducial.
Clustered components that have allowable transitions counts and transitions with
uniform widths are further considered for classification to determine
the binary code.

@@

\begin{figure}[h!]
\centering
  \includegraphics[width=0.8\linewidth]{training_data.pdf}
  \caption{Process to create training data for the fiducial pattern with binary
  code ``01''. The synthetic pattern is blurred along various orientations. Gabor
  filter is applied on the blurred pattern. Intensity profile is found along the
  first principal component of the clustered Gabor output. Same process is used
  to create training data for other fiducial patterns.}
  \label{fig:training_data}
\end{figure}

@@

\noindent\begin{figure*}
\begin{tabular}{c|c}
\begin{subfigure}[b]{0.65\linewidth}
\includegraphics[width=\linewidth]{recognition_rate.pdf}
\caption{Comparison of recognition rate of fiducials}
\label{fig:recognition_rate}
\end{subfigure}
&
\begin{subfigure}[b]{0.33\linewidth}
\begin{tabularx}{\textwidth}{|c|Y|}
\cline{1-2}
\footnotesize{Blur} & \footnotesize{Mean($\pm$std) error}  \\
\footnotesize{angle} & \footnotesize{(in pixels)}  \\
\cline{1-2}
\footnotesize{0} & \footnotesize{1.33 $\pm$ 0.40}  \\
\footnotesize{22} & \footnotesize{2.42 $\pm$ 1.05} \\
\footnotesize{45} & \footnotesize{1.46 $\pm$ 0.65}  \\
\footnotesize{67} & \footnotesize{2.39 $\pm$ 1.28}  \\
\footnotesize{90} & \footnotesize{1.83 $\pm$ 0.25}  \\ 
\cline{1-2}
\end{tabularx}
\caption{Center localization error}
\label{tab:blur_angle_center}
\end{subfigure}
\end{tabular}
\caption{(a): Comparison of recognition rate of the ARTag and our fiducials on
blur simulated data at various blur scales. It can be seen that except for the
fiducial with binary code ``00'', our fiducial patterns are recognized all times.
(b): Table showing the mean error and standard deviation of the localized fiducial center
versus the ground truth.  Error is computed for various blur angles over
different blur scales.}
\label{fig:simulated_blur}
\end{figure*}

@@

\begin{table}[t!]
\caption{Comparison of recognition rate of ARTag and our fiducials on real
data captured through AR Drone. Each row shows analysis of a test
dataset captured for our fiducial with different binary code embedded in it.
Each dataset has around 1000 frames captured representing roughly two minutes of video each.}
\centering
\begin{tabularx}{\linewidth}{|c|Y|Y|Y|Y|}
\cline{1-5}
\multirow{2}{*}{Test \#} & {Number of frames}
&{Binary Code} &\multicolumn{2}{c|}{Recognition Rate (\%)} \\
\cline{4-5} & & & ARTag & Our Fiducial \\\cline{1-5}
1 & 1205 & 00 &  65.6 & 86.5  \\ \cline{1-5}
2 & 1047 & 01 &  61.9 & 94.1  \\ \cline{1-5}
3 & 1102 & 10 &  62.4 & 92.74 \\ \cline{1-5}
4 & 1081 & 11 &  60.3 & 93.54  \\ \cline{1-5}
\end{tabularx}
\label{tab:recongition_accuracy}
\end{table}

In another setup we arranged our fiducials on four sides of a box and revolved
quadcopter around it. Later, we repeated the process by replacing our fiducial
by ARTags. Some intermediate frames are shown in Figure \ref{fig:setup}.

@@

We also compare our approach with tracking designed for blurred input scenes.
We have used four image sequences (consisting of around 1000 frames each), each
one containing a different blur resistant fiducial pattern to test the performance of
BLUT \cite{Wu:2011} and compare it with our result on the same image sequence.
From the images in the top rows of Figure
\ref{fig:BLUT_compare_00} and \ref{fig:BLUT_compare_01}, we can see that BLUT is
able to track the fiducial when the position of fiducial does not change too
much in successive frames. Also, it can be seen that once BLUT looses 
track of the fiducial, it is not able to recover. Since our approach detects the
code in each frame, large changes in the patterns position is not an issue.
Some of our detection results are also shown in Figure
\ref{fig:BLUT_compare_00} and \ref{fig:BLUT_compare_01}. (Results on other two
tags are shown in supplementary material.)

@@

We have demonstrated the effectiveness of our blur invariant fiducial both
synthetically and on real video clips captured from a quadcopter.   Our approach obtains
approximately 86-95\% recognition rate on fiducals in real scenes compared to existing codes
that are around 64\%.  We discuss some limitations of our approach in the
section.
%\noindent\textbf{Processing Time}~~
Currently, our processing time (0.3 seconds per frame) does not provide
real-time performance.  As such, we currently envision this will be used in an
offline manner for performing analysis of flight paths. Code profiling revealed
that the Gabor filtering along the eight directions takes most of the time
(0.03 -- 0.04 seconds per orientation).  Either an improved Gabor filter scheme
is required or an alternative strategy for detection is required.
%\noindent\textbf{False Negatives}~~
We found that sometimes our detection
algorithm fails to recognize our fiducial with binary code ``00'' when there is
large amount of blur.  The is because when there is too much blur, the innermost
ring's response in the Gabor output is too low and not detected properly.
This problem can be resolved by increasing the radius of innermost ring to
reduce the effect of blur on the innermost ring. 
%\noindent\textbf{Pose Estimation}~~
We note that other markers are able to give
full pose estimation after detection.  However, we are only able to reliably detect
the center point and therefore cannot estimate pose.  Of course, if four
markers were used in a known order, pose could be estimated. The current
resolution of onboard camera will be main hurdle to clear before we can
effectively use multiple markers in each scene.
%\noindent\textbf{Number of Fiducials}~~
Currently, compared to ARTag fiducial markers, we are able to generate less
number of fiducials. Many applications in robotics (e.g. quadcopter navigation)
may not require large number of fiducial markers as available by ARTag.  Most
of the time, it is sufficient to have 4-6 different fiducial patterns. Still,
we may be able to generate a larger number of fiducials by using color
backgrounds. This is an area of interest for further investigation.

>>>>

\usepackage{wacv}
\usepackage{times}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[font=small]{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{color}
\usepackage{url}
\usepackage{comment}
\usepackage{multirow}
\usepackage{tabularx}
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\setlength{\floatsep}{7pt plus 2pt minus 2pt}
\setlength{\textfloatsep}{5pt plus 2pt minus 2pt}
\setlength{\intextsep}{5pt plus 2pt minus 2pt}
% Include other packages here, before hyperref.

@@

\noindent\textbf{Contribution}~~To address these problems, we propose a
fiducial that is designed to be resistant to motion blur. Our design is based on
concentric circles as shown in Figure \ref{fig:teaser}-(c,d). The design is based
on the observation that motion blur from a quadcopter tends to be linear in
nature. As such, when our fiducial is blurred, there is no blur in the direction
perpendicular to the direction of motion.  This allows the signature of the
fiducial to remain intact in any direction. In addition, by using concentric
rings, we can treat the presence or absence of a ring as a bit, allowing us to
assign a code to the marker. Our experiments show that this design can
significantly outperform existing codes in the presence of motion blur. As far
as we are aware, this is the first work to propose a blur resistant marker.

@@

\begin{figure}
\includegraphics[width=\linewidth]{teaser.pdf}
\caption{This figure shows the experimental setup and comparison of
the output of ALVAR~\cite{alvar} (for detecting ARTag) versus our blur resistant
fiducial in a normal and blurred video frame. The scene contains two ARTags and our code.
(a) shows that the ARTag can be detected when the image has no blur.
(b) shows that ARTags are no longer detectable due to the blur.
(c-d) show the same images in (a) and (b) where our code is being detected.}
\label{fig:teaser}
\end{figure}

@@

\noindent{\textbf{Fiducials}}~Figure~\ref{fig:previous_work} shows several
examples of existing fiducial markers.  Many designs use a two dimensional
barcode inside a rectangular grid. One example of such a fiducial is from the
ARToolkit~\cite{ARToolkit02}, a well known toolkit used in many augmented
reality (AR) applications. Kato and Billinghurst~\cite{kato-artoolkit}
first demonstrated the use of ARToolkit in various augmented-reality-based applications.

Fiala~\cite{Fiala05} proposed a fiducial termed, ARTag, which is a
bi-tonal system consisting of a square border and an interior 6$\times$6 grid
of black or white cells. The improvement of the ARTag compared to the ARToolkit was
detection of corners instead of detection of lines to find possible pattern.
This proved to be more efficient than \cite{ARToolkit02} in terms of
recognition rate as well as the number of different patterns which can be
created.   The reliance on both line and corner detection, however, hampers recognition
under motion blur.

@@

\noindent{\textbf{Tracking}}~~Fiducial detection between successive video
frames can be considered as a tracking problem where the tracked object is
the fiducial.  There is a very large body of research dedicated to tracking and interested readers are referred to~\cite{Yilmaz:2006} for a good survey.

Most tracking methods (e.g.,~\cite{Ross:2008,Wu:2009,Perez02,Mei:2009}) assume
the image sequence to be blur free. In reality, however, the presence of motion blur in
a video sequence is often unavoidable. To this end, Wu et al.~\cite{Wu:2011}
proposed the Blur-driven Tracker (BLUT) framework for tracking motion-blurred targets. BLUT
is based on the observation that although motion blur degrades the visual
features of the target, it may also provide useful cues about the
movements to help tracking.

The BLUT framework successfully tracks a blurred target when there is uniform motion
and the position of the tracked object does not change drastically in successive
frames. However, as we will demonstrate in Section 4, the erratic motion from the quadcopters as well as the problem of dropped video frames is too problematic to be successively tracked.

@@

Here we examine the the performance of two
popular fiducials under motion blur.  Specifically we examine ARTags
~\cite{Fiala05} and PiTags~\cite{Pitag13} given their difference in geometric
design and the availability of API to develop applications to recognize the tags.
RUNE-tag~\cite{runetag11} currently does not provide access to the an
implementation to generate or recognize its tags.  The circular data
matrix~\cite{NaimarkF02} is available as a commercial product, however it
requires proprietary hardware to recognize the fiducial.

@@

\begin{figure}
\centering
  \includegraphics[width=.24\linewidth]{newconcentric_00.pdf}
  \includegraphics[width=.24\linewidth]{newconcentric_01.pdf}
  \includegraphics[width=.24\linewidth]{newconcentric_10.pdf}
  \includegraphics[width=.24\linewidth]{newconcentric_11.pdf}
  \caption{Two bit binary coded fiducials (from left to right: binary code 00,
  binary code 01, binary code 10, binary code 11.)}
  \label{fig:fiducials}
\end{figure}

Depending on which ring is present or absent, the resulting binary code will
change. The number of different patterns depends on the number of bits in the
binary code. For example, if the binary code has three bits, there will be a
maximum of three rings between marker-rings and we end up with eight
different patterns.  Figure \ref{fig:fiducials} shows two bit binary coded
fiducials.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{blur_direction.pdf}
\caption{Figure showing how change in blur direction changes the location of
unblurred linear pattern.}
\label{fig:blur_direction}
\end{figure}

@@

\begin{figure*}[ht!]
\centering
\includegraphics[width=0.8\linewidth]{overall_flow.pdf}
\caption{An overview of our algorithm to detect and classify fiducial.
The four step process includes (Step 1) Gabor filtering,
(Step 2) component clustering, (Step 3) PCA to determine the dominant direction,
and (Step 4) K-NN classification using training-data captured for each pattern.}
\label{fig:overall_flow}
\end{figure*}

@@

After finding the intensity profile, we examine it by
projecting the pixel intensities onto the PCA direction and record the
number and width of the white-to-black transitions.  If the number of transitions or the
width of the transitions are not consistent with what is allowable by
our code, we reject the clustered region as a potential fiducial.
Clustered components that have allowable transitions counts and transitions with
uniform widths are further considered for classification to determine
the binary code.

@@

\begin{figure}[h!]
\centering
  \includegraphics[width=0.95\linewidth]{training_data.pdf}
  \caption{Process to create training data for the fiducial pattern with binary
  code ``01''. The synthetic pattern is blurred along various orientations. Gabor
  filter is applied on the blurred pattern. Intensity profile is found along the
  first principal component of the clustered Gabor output. Same process is used
  to create training data for other fiducial patterns.}
  \label{fig:training_data}
\end{figure}

@@

\noindent\begin{figure*}
\centering
\begin{tabular}{c|c}
\begin{subfigure}[b]{0.45\linewidth}
\centering
\includegraphics[width=\linewidth]{recognition_rate.pdf}
\caption{Comparison of recognition rate of fiducials}
\label{fig:recognition_rate}
\end{subfigure}
&
\begin{subfigure}[b]{0.25\linewidth}
\centering
\begin{tabularx}{\textwidth}{|c|Y|}
\cline{1-2}
\footnotesize{Blur} & \footnotesize{Mean($\pm$std) error}  \\
\footnotesize{angle} & \footnotesize{(in pixels)}  \\
\cline{1-2}
\footnotesize{0} & \footnotesize{1.33 $\pm$ 0.40}  \\
\footnotesize{22} & \footnotesize{2.42 $\pm$ 1.05} \\
\footnotesize{45} & \footnotesize{1.46 $\pm$ 0.65}  \\
\footnotesize{67} & \footnotesize{2.39 $\pm$ 1.28}  \\
\footnotesize{90} & \footnotesize{1.83 $\pm$ 0.25}  \\
\cline{1-2}
\end{tabularx}
\caption{Center localization error}
\label{tab:blur_angle_center}
\end{subfigure}
\end{tabular}
\caption{(a): Comparison of recognition rate of the ARTag and our fiducials on
blur simulated data at various blur scales. It can be seen that except for the
fiducial with binary code ``00'', our fiducial patterns are recognized all times.
(b): Table showing the mean error and standard deviation of the localized fiducial center
versus the ground truth.  Error is computed for various blur angles over
different blur scales.}
\label{fig:simulated_blur}
\end{figure*}

@@

\begin{table}[t!]
\caption{Comparison of recognition rate of ARTag and our fiducials on real
data captured through AR Drone. Each row shows analysis of a test
dataset captured for our fiducial with different binary code embedded in it.
Each dataset has around 1000 frames captured representing roughly two minutes of video each.}
\centering
\begin{tabularx}{\linewidth}{|c|Y|Y|Y|Y|}
\cline{1-5}
\multirow{2}{*}{Test \#} & {Number}
&{Binary} &\multicolumn{2}{c|}{Recognition Rate (\%)} \\
\cline{4-5} & {of frames}& {Code}& ARTag & Our Fiducial \\\cline{1-5}
1 & 1205 & 00 &  65.6 & 86.5  \\ \cline{1-5}
2 & 1047 & 01 &  61.9 & 94.1  \\ \cline{1-5}
3 & 1102 & 10 &  62.4 & 92.74 \\ \cline{1-5}
4 & 1081 & 11 &  60.3 & 93.54  \\ \cline{1-5}
\end{tabularx}
\label{tab:recongition_accuracy}
\end{table}

In another setup we arranged our fiducials on four sides of a box and revolved
the quadcopter around the box. Later, we repeated the process by replacing our fiducial
by ARTags. Figure \ref{fig:setup} some frames from this sequence as well as the performance of both fiducials. The overall detection rate of our tags is 90\% while that of ARTags is only around 60\%.

@@

We also compare our approach with tracking designed for blurred input scenes.
We have used four image sequences (consisting of around 1000 frames each), each
one containing a different blur resistant fiducial pattern to test the performance of
BLUT \cite{Wu:2011} and compare it with our result on the same image sequence.
From the images in the top rows of Figure
\ref{fig:BLUT_compare_00} and \ref{fig:BLUT_compare_01}, we can see that BLUT is
able to track the fiducial when the position of fiducial does not change too
much in successive frames. Also, it can be seen that once BLUT looses
track of the fiducial, it is not able to recover. Since our approach detects the
code in each frame, large changes in the patterns position is not an issue.
Some of our detection results are also shown in Figure
\ref{fig:BLUT_compare_00} and \ref{fig:BLUT_compare_01}.

@@

We have demonstrated the effectiveness of our blur invariant fiducial both
synthetically and on real video clips captured from a quadcopter.   Our approach obtains
approximately 86-95\% recognition rate on fiducals in real scenes compared to existing codes
that are around 64\%.  We discuss some limitations of our approach in the
section.
%\noindent\textbf{Processing Time}~~
Our  current processing time (0.3 seconds per frame) does not provide
real-time performance.  As such, we envision this will be used in an
offline manner for performing analysis of flight paths. Code profiling revealed
that the Gabor filtering along the eight directions takes most of the time
(0.03 -- 0.04 seconds per orientation).  Either an improved Gabor filter scheme
is required or an alternative strategy for detection is required.
%\noindent\textbf{False Negatives}~~
We found that sometimes our detection
algorithm fails to recognize our fiducial with binary code ``00'' when there is
large amount of blur.  The is because when there is too much blur, the innermost
ring's response in the Gabor output is too low and not detected properly.
This problem can be resolved by increasing the radius of innermost ring to
reduce the effect of blur on the innermost ring.
%\noindent\textbf{Pose Estimation}~~
We note that other markers are able to give
full pose estimation after detection.  However, we are only able to reliably detect
the center point and therefore cannot estimate pose.  Of course, if four
markers were used in a known order, pose could be estimated. The current
resolution of onboard camera will be main hurdle to clear before we can
effectively use multiple markers in each scene.
%\noindent\textbf{Number of Fiducials}~~
Currently, compared to ARTag fiducial markers, we are able to generate less
number of fiducials. Many applications in robotics (e.g. quadcopter navigation)
may not require large number of fiducial markers as available by ARTag.  Most
of the time, it is sufficient to have 4-6 different fiducial patterns. Still,
we may be able to generate a larger number of fiducials by using color
backgrounds. This is an area of interest for further investigation.
